---
title: "AE 03: Duke Forest + data visualization"
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

## Exercise 1

Suppose you're helping some family friends who are looking to buy a house in Duke Forest.
As they browse Zillow listings, they realize some houses have garages and others don't, and they wonder: **Does having a garage make a difference?**

Luckily, you can help them answer this question with data visualization!

-   Make histograms of the prices of houses in Duke Forest based on whether they have a garage.
    -   In order to do this, you will first need to create a new variable called `garage` (with levels `"Garage"` and `"No garage"`).
    -   Below is the code for creating this new variable. Here, we `mutate()` the `duke_forest` data frame to add a new variable called `garage` which takes the value `"Garage"` if the text string `"Garage"` is detected in the `parking` variable and takes the test string `"No garage"` if not.

```{r}
#| label: garage
#| eval: false

duke_forest <- duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage"))

```

-   Then, facet by `garage` and use different colors for the two facets.
-   Choose an appropriate binwidth and decide whether a legend is needed, and turn it off if not.
-   Include informative title and axis labels.
-   Finally, include a brief (2-3 sentence) narrative comparing the distributions of prices of Duke Forest houses that do and don't have garages. Your narrative should touch on whether having a garage "makes a difference" in terms of the price of the house.

```{r}
#| label: price-garage

ggplot(duke_forest, aes( x = price, fill = garage, colour = garage)) +
  geom_histogram(binwidth = 100000)+
  facet_wrap(~garage) +
  labs(
    x = "Price",
    y = "Number of Houses",
    title = "House Prices by Garage Presence"
  )

```

Interpretations of the histogram allow for the conclusion that houses without garages have the potential to be more appealing to the consumer.
This is because the distribution of homes without garages lies on the lower end of the price scale (\$400,000 and \$700,000), with more frequency of houses.
Comparatively, there appear to be fewer houses with garages in Duke Forest and, of the houses, they are on the higher end of the price scale (\$600,000 to \$900,000).

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

## Exercise 2

It's expected that within any given marker larger houses will be priced higher.
It's also expected that the age of the house will have an effect on the price.
However in some markets new houses might be more expensive while in others new construction might mean "no character" and hence be less expensive.
So your family friends ask: "In Duke Forest, do houses that are bigger and more expensive tend to be newer ones than those that are smaller and cheaper?"

Once again, data visualization skills to the rescue!

-   Create a scatter plot to exploring the relationship between `price` and `area`, conditioning for `year_built`.
-   Use `geom_smooth()` with the argument `se = FALSE` to add a smooth curve fit to the data and color the points by `year_built`.
-   Include informative title, axis, and legend labels.
-   Discuss each of the following claims (1-2 sentences per claim). Your discussion should touch on specific things you observe in your plot as evidence for or against the claims.
    -   Claim 1: Larger houses are priced higher.
    -   Claim 2: Newer houses are priced higher.
    -   Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

```{r}
#| label: price-area
ggplot(duke_forest, aes(x = area, y = price, colour = year_built))+
  geom_point()+
  geom_smooth(se = FALSE)+
  labs(
     x = "Area",
    y = "Price" ,
    title = "House Price vs. Area by Year Built"
  )
```

-   Claim 1: Based on the trendline produce, we see and general positive relationship between Area and Price. This we can conclude a correlation that an increase in area for a house lead to the house being priced higher.
-   Claim 2: The plots show significant overlap in price for older house and new houses in the \$400000 - \$700000 range, therefore it cannot be concluded that year of build contributes significantly to price.
-   Claim 3: Considering plots that represent the newest built houses, it shows that they stay within a limited range near the middle for both area and price. Thus the data is unable to confirm the claim.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::
